---
title: "Quick start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r child = 'Contents.Rmd'}
```


The simplest way to run tugHall is just run function:

    simulation()

Also, the code of the simulation is presented in the pipeline 'pipeline_simulation.R' (that can be obtained by command copy_pipelines() )


This function makes several changes in the working directory:

- copy /Input/ folder to your working directory;

- create /Output/ folder in your working directory;

- define all the global parameters;

- define all the names of files for input and output;

- define genes location info from the file;

- define hallmarks from the file;

- define compaction factors from the file;

- define initial clones;

- run the simulation (please, wait for a while);

- after finishing of simulation, it calculates VAF and order of genes dysfunction;

- plot given results.


