---
title: "Preparation of simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preparation of simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Table of Contents

[User Guide content](A_User_Guide.html)

[Theory of the model](B_Theory.html)

[Quick start guide](C_Quick_start.html)

[Simulation preparation](D_Simulation.html)

[Input data](E_Input.html)

[Output data](F_Output.html)

[Post simulation analisys](G_Analisys.html)

[VAF calculation](H_VAF.html)

[Figures](I_Figures.html)

[Relation to experimental measurements](J_Relation.html)

[Drug intervention](K_Drug_intervention.html)

[Versions](L_Versions.html)

[References](M_References.html)


---

# Standart simulation         
        
Standard simulation includes several steps:

1. Preparation of Input and Output folders from working directory

2. Preparation of input files and parameters

3. Run a simulation

4. Post simulation data processing 

5. Plot the results of simulation

6. Split and plot original simulation separately for drivers and passengers:



### 1. Preparation of Input and Output folders from working directory

Load the tugHall.3 simulation package:

```{r, echo=TRUE}
library(tugHall.3)
```
            
        
Definen number of **digits** in the numeric format for tugHall.3 environment:

```{r, echo=TRUE}
pck.env$digits = 6
```


Attach packages from import list
            
```{r, echo=TRUE}
check_packages()
```        


### 2. Preparation of input files and parameters

Firstly, we need to copy all the input files to Input folder by calling the function,
(make sure that **working directory** is in a proper place):

        copy_files_to_Input( files = c( 'CCDS.current.txt', 'CF.txt',
                                        'cloneinit.txt','gene_hallmarks.txt',
                                        'gene_map.txt','parameters.txt' ) ,
                                         dir = 'Input' )


            
Next define names of all the files to input and output:

        define_files_names()

Define location of all the genes of interest:

        define_gene_location()

Next check/view the default gene_map from the CCDS.current.txt file. Here the model will show the default gene list from the file. If you need to change the gene list change/define the gene "CCDS_ID" in the file:

        pck.env$gene_map

Define parameters from './Input/parameters.txt', if it is necessary, please, change it directly:


        define_parameters( read_fl = TRUE , file_name = './Input/parameters.txt' )


If compaction factors are used in simulation, please, change and define them:

        define_compaction_factor( read_fl = TRUE , file_name = './Input/CF.txt' )

Before simulation, check that the previous results are deleted and delete, if not. Also check the input parameters and compaction factors before simulation:

        check_previous_data( )

      

### 3. Run a simulation

Start **simulation**, please, wait for a while ... or see Sim_monitoring.txt file in working folder:

        smlt = model()
        
Save point mutations and CNA to the files in Output folder:


        write_pnt_clones( pck.env$pnt_clones, file_out = 'Output/point_mutations.txt' )
        write_pnt_clones( pck.env$cna_clones, file_out = 'Output/CNA_mutations.txt' )


### 4. Post simulation data processing 

Get data of the last simulation from cloneout file and gene file:

        dtst = get_flow_data( pck.env$cloneoutfile, pck.env$genefile )
                    pck.env$data_avg   =  dtst$data_avg
                    pck.env$data_flow  =  dtst$data_flow
                    pck.env$time_max   =  dtst$time_max
                    pck.env$data_last  =  dtst$data_last
                    cna_mut = dtst$cna_mut
                    pnt_mut   =  dtst$pnt_mut
                    pnt_mut_B =  dtst$pnt_mut_B

            
Please check the Onco  and hall objects after simulation and verify it with the input data


        pck.env$onco
        pck.env$hall
        
Also get VAF data and save them into file 'Output/VAF_data.txt':


        pck.env$vf = get_VAF( pnt_mut, pck.env$data_last )
        pck.env$VAF  =  get_rho_VAF( vf = pck.env$vf, rho = c( 0.0, 0.1, 0.2, 0.5, 0.7, 0.9 ) ,
                                     file_name = './Output/VAF.txt' )

     
If it necessary, get order of genes dysfunction for all clones:


        pck.env$rdr_dysf  =  get_order_of_genes_dysfunction( 
                        pnt_mut = pnt_mut,
                        pck.env$data_last, cna_mut,
                        file_name ='./Output/order_genes_dysfunction.txt' )
        
### 5. Plot the results of simulation

To visualize results of a simulation one may plot them using special functions:

        rdr_dysf = tugHall_dataset$rdr_dysf
        plot_order_dysfunction( rdr_dysf , pos = c(5,800), logscale = 'y', cex = 1. )
        data_avg = tugHall_dataset$data_avg
        plot_average_simulation_data( data_avg, pck.env$time_max )
        

Plot the main clones

        data_flow = tugHall_dataset$data_flow
        plot_clone_evolution( data_flow, threshold = c(0.01, 1 ), lwd = 2.0,
                                  hue = c(" ", "random", "red", "orange", "yellow",
                                          "green", "blue", "purple", "pink", "monochrome")[1],
                                  luminosity = c(" ", "random", "light", "bright", "dark")[4],
                                  yr = NA , add_initial = TRUE, log_scale = TRUE )


Plot the minor clones but large amount of them


        data_flow = tugHall_dataset$data_flow
        plot_clone_evolution( data_flow, threshold = c(0.0, 0.01), lwd = 2.0,
                                  hue = c(" ", "random", "red", "orange", "yellow",
                                          "green", "blue", "purple", "pink", "monochrome")[1],
                                  luminosity = c(" ", "random", "light", "bright", "dark")[4],
                                  yr = NA , add_initial = FALSE, log_scale = TRUE )


Get/store the results from the simulation 

        res  =  get_tugHall.Environment()
                    vf_init  =  pck.env$vf


Plot the Variant Allele Frequency (VAF) data for normal, primary and metastatic cells

        VAF = tugHall_dataset$VAF
        plot_VAF( VAF = VAF, rho = 0 , violin = FALSE )


### Split and plot original simulation separately for drivers and passengers:

Split the VAF from original simulation results for drivers and passengers:

         vf_drivers  =  vf_init[ which(  vf_init$MalfunctionedByPointMut ), ]
                    vf_pass     =  vf_init[ which( !vf_init$MalfunctionedByPointMut ), ]
                    VAF_drivers  =  get_rho_VAF( vf = vf_drivers, save_to_file = FALSE )
                    VAF_pass  =  get_rho_VAF( vf = vf_pass, save_to_file = FALSE )

Plot the Variant Allele Frequency (VAF) data of drivers and passengers seperately for normal, primary and metastatic cells

        if (length(VAF_drivers) > 0 ) 
                plot_VAF( VAF = VAF_drivers, rho = 0 , violin = FALSE )
        if (length(VAF_pass) > 0 ) 
                plot_VAF( VAF = VAF_pass,    rho = 0 , violin = FALSE )



# Parallel simulations

To be done

# Simulations with different parameters

To be done 

# Simulation started from saved results of previous simulation 

To be done 


