---
title: "Relation to experimental measurements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Relation to experimental measurements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r child = 'Contents.Rmd'}
```


We here list variables processed from the **tugHall.3** outputs that are related to experimental measurements. 


|  Variables processed from the simulator outputs  | Relation to experimental measurements  |
|:---|:---|
| Number of cells | Observed tumor size. $10^9$ cells correspond to the tumor tissue diameter of 1 cm. $10^{12}$ cells correspond to that of 10 cm. $10^{12}-10^{13}$ cells correspond to lethal burden. See Friberg and Mattson, Journal of Surgical Oncology, 1997. |
| VAF | VAF calculated from sequence reads in the next-generation sequencer (NGS) under the assumption of tumor purity as external parameter, for details, please, see [VAF calculation](H_VAF.html). |
| Mutation number per base-pairs | Tumor mutation burden calculated from NGS data. |
| Number of clones | Number of clones estimated from NGS data by computational tools such as SciClone ([Miller et al, PLOS Computational Biology, 2014](https://doi.org/10.1371/journal.pcbi.1003665)) and SubClonalSelection ([Williams et al, Nature Genetics, 2018](https://doi.org/10.1038/s41588-018-0128-6)). |
| <img width=200/> | <img width=440/> |




